\Large \bf DRAFT: Direct-style process creation on Linux

### 2020-05-23 research/paper/paper.tex
paper!!

\usepackage{tikz}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{pifont}
\usepackage{comment}

@@

\section{Introduction}\label{introduction}
Existing process creation interfaces can be divided into three categories:
Fork-style, spawn-style, and direct-style.

Fork-style (such as \texttt{fork}) and spawn-style (such as \texttt{posix\_spawn}) are the most popular,
but have a number of disadvantages.
Correct, performant use of fork-style places significant constraints on the calling process,
and also has a complex programming model.
Spawn-style lifts many of these constraints, and is simpler to program against,
but is fundamentally limited in the kinds of processes it can create.
With both fork-style and spawn-style,
it is difficult to
branch based on results or errors encountered while initializing the new process:
fork-style requires IPC if it wishes to communicate errors back to the rest of the program,
and robust indication of errors in spawn-style requires substantial expansion of the spawn interface.

@@

Our contribution is an implementation of direct-style process creation for Linux,
making use of our work on the \texttt{rsyscall} project.
The \texttt{rsyscall} project develops language-specific ``process-independent'' system call libraries for Linux,
which replace libc system call wrappers,
and allow a program to perform cross-process system calls.
\texttt{rsyscall} works entirely in userspace
and can be deployed on existing Linux systems without kernel modifications.
\texttt{rsyscall} is currently focused on supporting, and is implemented in, Python,
but can be ported to other languages;
the examples we show in this paper will be in Python,
but generalize easily to other languages.

Our direct-style process creation implementation uses \texttt{clone} to create a child process in an embryonic state
which can then be manipulated from user code in the parent process using syscalls with \texttt{rsyscall}.
A direct-style \texttt{clone} has all the advantages of direct-style process creation
over Linux's fork-style \texttt{fork} and spawn-style \texttt{posix\_spawn}:
\begin{compactitem}
\item
  Processes can be created correctly and performantly from any parent process,
  with a straightforward, familiar API.
\item
  Processes can be created with the full set of Linux features.
\item
  Syscall results are reported individually to the parent process.
  % rather than batched together as with spawn-style or manually reported via IPC as with fork-style.
\end{compactitem}

We've implemented a variety of useful applications using direct-style process creation on Linux,
which we'll demonstrate in this paper.
Direct-style \texttt{clone} has minimal performance cost in normal situations,
and can outperform \texttt{fork} in some cases.
We have deployed direct-style process creation in production internally at \twosigma.

Our implementation is distributed as a feature of \texttt{rsyscall},
and is immediately deployable, entirely in userspace, on recent Linux kernels.
\texttt{rsyscall} is open source and available from
[Github url].
% \url{https://github.com/catern/rsyscall}.

This paper is organized in several sections.
\begin{compactitem}
\item Section \ref{background} provides background on fork-style, spawn-style, and direct-style process creation,
and considers their advantages and disadvantages in more depth.
\item Section \ref{overview} discusses the design of the API,
including our usage of \texttt{rsyscall}.
\item Section \ref{examples} shows examples of direct-style process creation in Python on Linux,
including use of process features which are difficult to use on Linux with \texttt{fork} or \texttt{posix\_spawn}.
\item Section \ref{implementation} explores our implementation in depth,
and discusses several difficult process-creation related details of \texttt{rsyscall}.'s implementation.
\item Section \ref{evaluation} evaluates our results using both quantitative and qualitative approaches.
\item Section \ref{related_work} examines other work which also touches on the topics covered by this paper.
\item Section \ref{future_work} lists several directions for future work
related to direct-style process creation and to \texttt{rsyscall}.
\item Section \ref{conclusions} briefly outlines our conclusions about direct-style and the future of process creation.
\end{compactitem}
\section{Background on process-creation}\label{background}
As mentioned in the introduction,
on most operating systems today,
the available interfaces for process creation
are either "fork-style" or "spawn-style".
A few operating systems have a third style, which we refer to as "direct-style".

\begin{table*}
\begin{tabular}{l|l|l|l}
% BEGIN RECEIVE ORGTBL compare
 & fork-style & spawn-style & direct-style\\
\hline
Works with any caller & \tbad No threads or large memory & \tgood Yes & \tgood Yes\\
Programming model & \tbad Complex (returns twice) & \tgood Simple (single call) & \tgood Simple (imperative)\\
Maximally powerful & \tgood Yes, can call any syscall & \tbad No, limited interface & \tgood Can call any syscall\\
Error handling & \tbad Poor, requires IPC & \tbad Poor, not fine-grained & \tgood Reports individual errors\\
Non-copied attributes & Mutated by code in child & Set by arguments & Mutated by code in parent\\
% END RECEIVE ORGTBL compare
\end{tabular}
\caption{Features of fork-style vs spawn-style vs direct-style}
\label{tab:ease}
\end{table*}
\begin{comment}
#+ORGTBL: SEND compare orgtbl-to-latex :splice t
|                       | fork-style                       | spawn-style                  | direct-style                     |
|-----------------------+----------------------------------+------------------------------+----------------------------------|
| Works with any caller | \tbad No threads or large memory | \tgood Yes                   | \tgood Yes                       |
| Programming model     | \tbad Complex (returns twice)    | \tgood Simple (single call)  | \tgood Simple (imperative)       |
| Maximally powerful    | \tgood Yes, can call any syscall | \tbad No, limited interface  | \tgood Can call any syscall      |
| Error handling        | \tbad Poor, requires IPC         | \tbad Poor, not fine-grained | \tgood Reports individual errors |
| Non-copied attributes | Mutated by code in child         | Set by arguments             | Mutated by code in parent        |
% $
\end{comment}
\subsection{Fork-style}
By far the most well-known fork-style interface is Unix's fork.\cite{forkhist}
With a fork-style interface,
a process, called the parent process, calls \texttt{fork},
and a copy of that process is made,
called the child process,
differing only in a few attributes such as process id.
The same program continues executing in both processes,
with a different return value from fork to allow it to distinguish whether it is running in the parent or the child.
Typically, the program will branch based on the return value
and, if running in the child,
call various system calls to modify the state of the new process.

Fork has a number of flaws,
and much ink has been spilled detailing them.
We'll cover the most egregious briefly here,
but see Baumann 2019 \cite{forkroad} for a more detailed evaluation.

Fork, when called successfully, returns twice,
creating two different processes which execute concurrently.
This immediately makes fork difficult to integrate;
for example, errors can't be easily communicated between the two processes.
If there is an error in the initialization of the child process,
the part of the program running in the child process
needs to use some form of inter-process communication (IPC) to communicate it back to the parent process.
The child process can run arbitrary code to make decisions about its initialization,
but that code needs to use IPC if it wishes to communicate with the rest of the program,
which is still in the parent process.

Fork can have poor performance when called from processes with many memory mappings.\cite{forkroad}
Fork copies many attributes about the parent process when creating the child process,
including setting up copy-on-write memory mappings in the child process.
This becomes slower as the parent process has more memory mappings,
eventually taking a significant amount of time.
This copying is required to robustly implement fork's model,
where the same program continues executing without changes in both the parent process and child process;
the copy of the program running in the child process might access any memory at any time.

Multi-threaded programs generally cannot use fork safely.
Typical Unix fork implementations duplicate only the thread calling fork from the parent process to the child process.
In a multi-threaded program, this can cause various issues;
for example, another thread might be holding a memory allocator lock at the time of the fork,
which in the child process will never be unlocked,
causing the child process to deadlock if it tries to allocate memory.
Some thread libraries provide partial mitigations for this issue,
but it's up to user code to make use of those mitigations.\cite{pthread_atfork}
In combination with fork's poor performance in large-memory programs,
a user of fork must think carefully
about the characteristics of the process from which fork is being called.
\subsection{Spawn-style}
In a spawn-style interface,
details about the new process are provided up-front as arguments to some function
which creates the new process all at once.
The most well known spawn-style interfaces are Unix's \texttt{posix\_spawn} \cite{posix_spawn}
and Windows \texttt{CreateProcess} \cite{create_process};
the Unix shell also serves as a spawn-style process creation interface,
especially when used with Bernstein chaining\cite{chainloading}.

Spawn-style interfaces typically still transparently copy some details from the parent process;
for example, various security contexts,
and any other process attribute which is not explicitly specified in the arguments to the spawn interface.
The key difference between fork-style and spawn-style is not how much they copy,
it is how the attributes which are not copied are specified:
by mutation from code running in the new process in the case of fork-style,
or by explicit argument passing in the case of spawn-style.

Spawn-style interfaces lift the constraints on the calling process that fork-style interfaces impose.
Spawn-style interfaces don't run user code in the new process during initialization.
This means there is no need to copy memory,
so performance is good when a spawn-style interface is used from a process with many memory mappings.
Likewise, there are no inherent issues when a spawn-style interface is used from a multi-threaded program,
though operating systems, such as Linux,
which implement spawn-style interfaces on top of \texttt{fork} can still have threading bugs.

The arguments that can be provided to a spawn-style process creation syscall
do not cover all the possible attributes that one might want to set for the new process.
Most systems have a large number of syscalls which can mutate the state of a process during its lifetime;
for a spawn-style interface to work in all scenarios,
all those possible mutations must be reproduced in the interface.
For example, the \texttt{posix\_spawn} function provided by glibc does not support creating processes in new namespaces,
as is required for container functionality on Linux.

Spawn-style process creation
essentially creates a new domain-specific language for specifying the attributes of the process.
Fork-style allows existing system calls,
with which programmers are already familiar,
to be used to set up the new process.
Spawn-style requires some form of declarative description of the new process,
which must be learned in addition to the normal system call interface.

Spawn-style process creation also does not allow for conditional logic during the setup.
If the setup of the new process encounters an error at some point,
the only option is to return from the entire spawn call with an error.
Such errors returned from spawn-style calls
are typically much less informative
than the errors returned by the syscalls which directly mutate the process attributes.
Other forms of conditional logic are also impossible in spawn-style;
one modification to the process cannot depend on the result of some other modification.
\subsection{Direct-style}
A few academic operating systems, such as KeyKOS \cite{keykos}, seL4 \cite{sel4}
and some others \cite{exokernel} \cite{fuschia} \cite{singularity},
use another style of process creation.
We use the new term "direct-style" to refer to this,
since we know of no pre-existing concise and generic term for this style of process creation.

@@

However, we believe that a natural port of direct-style process creation to Linux
provides an interface that is just as simple to use as fork-style or spawn-style on Linux,
without their disadvantages.
Such an interface, like fork-style and spawn-style,
would implicitly copy some attributes to the new process,
rather than being capability-oriented and requiring all attributes to be specified explicitly.
This is the kind of direct-style interface that we contribute in this paper.
\section{Overview of API}\label{overview}
In most operating system APIs, including the ones available on Linux,
when a program makes a syscall, it implicitly operates on the current process.
To perform direct-style process creation on Linux,
we need an API where we instead explicitly specify in which process we want to make a syscall.

\texttt{rsyscall} provides this API.
The \texttt{rsyscall} project develops
language-specific, object-capability-model, low-abstraction libraries for the Linux syscall interface,
bypassing libc.

To perform direct-style process creation,
we use \texttt{rsyscall} to call \texttt{clone} on the ``local'' process to create a new child process.
We call various syscalls in the child process to mutate it until it reaches the desired state,
at which point we can call \texttt{execve} in the child process to start it running on its own.
After that, the process is no longer under our control, and it functions like any other child process.
The child process can be monitored using normal Linux child monitoring syscalls,
such as \texttt{waitid}.

\lstinputlisting[float=ht,language=Python,label={basic},caption={Creating a new process, changing CWD, and execing}]
{examples/basic/direct.py}
Listing \ref{basic} shows this in action in Python.
We create a new child process using \texttt{clone} on the local process \verb|local|,
and receive back a capability to control it.
We call \texttt{chdir} in the child process to change its working directory to \texttt{/dev}.
Then we call \texttt{execve} in the child,
passing the path of the \texttt{cat} executable,
the arguments to the new executable,
and the unchanged set of environment variables.
The name of the executable is passed as the first argument,
as is typical in Unix.
Calling \texttt{execve} consumes the capability,
and later use of this child process capability will fail with an exception.

The behavior of system calls made on a remote process through \texttt{rsyscall}
is identical to the behavior of system calls made implicitly on the current process through libc.
This is an API that systems programmers are already familiar with,
and which is already thoroughly documented through Linux system call manpages.
The novelty is constrained to the newly explicit process capabilities.

The lack of novelty in the API is important not only for the learning curve,
but also for power.
Since we are using conventional Linux system calls,
just in a cross-process way,
direct-style \texttt{clone} will always be at least as powerful as \texttt{fork}.
\section{Examples}\label{examples}
In this section,
we'll demonstrate direct-style process creation
by demonstrating several advanced uses of processes
using more sophisticated Linux features.
For concision, we assume in these examples that we are running with sufficient permissions in these examples,
but these examples can all be run without privileges with appropriate use of \verb|CLONE_NEWUSER|.

@@

\lstinputlisting[float=*ht,language=Python,label={fds},caption={Passing down FDs}]
{examples/fds/direct.py}
In Listing \ref{fds},
we first open a file with read-write permission in the local process.
Then we create a child process,
which inherits all file descriptors from its parent.
We indicate that we want to operate on the child's inherited copy of the file descriptor with \verb|child.inherit_fd|.
Then we disable \texttt{CLOEXEC} on the child's file descriptor,
which on Linux we can do by clearing the FD flags with \verb|fcntl(fd, F_SETFD, 0)|.
Finally, we execute some executable in the child process,
passing the file descriptor number as an argument;
since we disabled \texttt{CLOEXEC} on the file descriptor,
the executable will be able to use that file descriptor through that file descriptor number.
Unlike in listing \ref{basic}, we call the \texttt{execv} wrapper rather than \texttt{execve},
so that we don't need to pass the (unmodified) environment to every \texttt{exec} call in these examples.
\subsection{Non-shared-memory concurrency}
Processes run concurrently,
which allows exploiting the parallelism of the underlying hardware.
Since processes don't share memory,
they can provide a less complex parallel programming environment
than shared-memory thread-based approaches.

>>>>

\usepackage{tikz}
\usepackage{dblfloatfix}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{pifont}
\usepackage{comment}

@@

\section{Introduction}\label{introduction}
\begin{table*}
\begin{tabular}{l|l|l|l}
% BEGIN RECEIVE ORGTBL compare
 & fork & spawn & direct-style\\
\hline
Requirements on parent process & \tbad Single-threaded, small memory & \tgood None & \tgood None\\
Programming model & \tbad Complex (returns twice) & \tgood Simple (single call) & \tgood Simple (imperative)\\
Maximally powerful & \tgood Yes, can call any syscall & \tbad No, limited interface & \tgood Yes, can call any syscall\\
Reporting of results & \tbad Requires IPC & \tbad Not fine-grained & \tgood From individual child syscalls\\
Non-inherited attributes & Mutated by code in child & Set by arguments & Mutated by code in parent\\
% END RECEIVE ORGTBL compare
\end{tabular}
\caption{Features of fork-style vs spawn-style vs direct-style}
\label{tab:styles}
\end{table*}
\begin{comment}
#+ORGTBL: SEND compare orgtbl-to-latex :splice t
|                                | fork                                | spawn                       | direct-style                          |
|--------------------------------+-------------------------------------+-----------------------------+---------------------------------------|
| Requirements on parent process | \tbad Single-threaded, small memory | \tgood None                 | \tgood None                           |
| Programming model              | \tbad Complex (returns twice)       | \tgood Simple (single call) | \tgood Simple (imperative)            |
| Maximally powerful             | \tgood Yes, can call any syscall    | \tbad No, limited interface | \tgood Yes, can call any syscall      |
| Reporting of results           | \tbad Requires IPC                  | \tbad Not fine-grained      | \tgood From individual child syscalls |
| Non-inherited attributes       | Mutated by code in child            | Set by arguments            | Mutated by code in parent             |
% $
\end{comment}
The most well-known process creation interface on Linux, and Unix in general, is \texttt{fork}.
When a parent process creates a child process through \texttt{fork},
or other similar fork-style interface,
both processes execute concurrently,
and the child process typically inherits all but a few attributes from the parent process.\cite{tlpi}
The child process can call arbitrary system calls
to mutate the new process until it reaches the desired state,
typically ultimately calling \texttt{exec}.

A user of fork must think carefully about the characteristics of the process from which fork is being called.
Fork can have poor performance when called from processes with many memory mappings.\cite{forkroad}
This can be mitigated to some degree, but inflicts further complexity on the caller.\cite{vfork}\cite{vfork_dangerous}
Multi-threaded programs generally cannot use fork safely.\cite{forkroad}\cite{tlpi}
Some thread libraries provide partial mitigations for multi-threading issues,
but it's up to user code to make use of those mitigations.\cite{pthread_atfork}

Since the child process runs concurrently with the parent process,
communicating events in the child back to the parent requires some form of IPC,
complicating process creation.\cite{posix_spawn_error_pipe}
Even the basic interface of fork - one function call which returns twice - is unusual;
such functions are generally considered to be complex, advanced topics in other settings.
\cite{callcc_fork}\cite{continuations}\cite{bashcc}

In a spawn-style interface,
details about the new process are provided up-front as arguments to some function
which creates the new process all at once;
\texttt{posix\_spawn} is a typical example of this style.\cite{posix_spawn}\cite{create_process}\cite{chainloading}
Other details, such as security context, are inherited from the parent process.\cite{tlpi}\cite{create_process}
Spawn-style APIs
can be called from any parent process without concerns about memory usage or multi-threading,
when used correctly.
\cite{forkroad}\cite{posix_spawn_error_pipe}

Unlike \texttt{fork}, a spawn-style interface provides an API for process creation
which is independent from existing system calls.\cite{spawn_h}\cite{subprocess_run}
Spawn-style interfaces therefore often lack support for some features,
and they also require learning a new interface distinct from the normal system call API,
both of which encourage systems programmers to avoid them.\cite{tlpi}\cite{posix_spawn_lwn}
Also, spawn-style interfaces, by their nature as a single call,
don't allow programs to branch on the results of individual modifications to the new process,
and have complex, or substantially less information, error reporting for such modifications.
\cite{posix_spawn}\cite{posix_spawn_error_pipe}

Table \ref{tab:styles} summarizes the differences between fork-style and spawn-style.
Each has its own advantages and disadvantages.

A few academic operating systems - KeyKOS, seL4, and some others, for example -
use another style of process creation.\cite{keykos}\cite{sel4}\cite{exokernel}\cite{fuschia}\cite{singularity}
We use the new term "direct-style" to refer to this,
since we know of no pre-existing generic term for this style of process creation.

@@

We'll discuss each style in more detail in section \ref{background}.

To perform direct-style process creation on Linux,
we need an API where we explicitly specify in which process we want to make a syscall.
This differs from the normal mode of operation on Linux,
where programs making syscalls implicitly operate on the current process.

Our contribution in this paper is this API, through the \texttt{rsyscall} library.
The \texttt{rsyscall} project develops
language-specific, object-capability-model, low-abstraction libraries for the Linux syscall interface,
bypassing libc.
\texttt{rsyscall} is currently focused on supporting Python,
but can be ported to other languages;
the examples we show in this paper will be in Python,
but generalize easily to other languages.
\texttt{rsyscall} is open source and available from
[Github url].

To perform direct-style process creation,
we use \texttt{rsyscall} to call \texttt{clone} on the ``local'' process to create a new child process.
We call various syscalls in the child process to mutate it until it reaches the desired state,
at which point we can call \texttt{exec} in the child process to start it running on its own.
After that, the process is no longer under our control, and it functions like any other child process.
The child process can be monitored using normal Linux child monitoring syscalls,
such as \texttt{waitid}.

These child process capabilities are created and managed in userspace,
by launching new processes running a syscall server,
which receives syscall requests, performs the syscall, and returns the response.
No kernel modifications are required, and \texttt{rsyscall} is immediately deployable on recent Linux kernel.
We discuss the implementation in depth in section \ref{implementation},
and examine several difficult process-creation related details.

\lstinputlisting[float=h,language=Python,label={basic},caption={Create new process, change CWD, and exec}]
{examples/basic/direct.py}
Listing \ref{basic} shows a program using \texttt{rsyscall}.
We create a new child process using \texttt{clone} on the local process \verb|local|,
and receive back a capability to control it.
We call \texttt{chdir} in the child process to change its working directory to \texttt{/dev}.
Then we call \texttt{execv} in the child,
passing the path of the \texttt{cat} executable
and the arguments to the new executable.
The name of the executable is passed as the first argument,
as is typical in Unix.
Calling \texttt{execv} consumes the capability,
and later use of this child process capability will fail with an exception.

The behavior of system calls made on a remote process through \texttt{rsyscall}
is identical to the behavior of system calls made implicitly on the current process through libc.
This is an API that systems programmers are already familiar with,
and which is already thoroughly documented through Linux system call manpages.
The novelty is constrained to the newly explicit process capabilities.
Besides flattening the learning curve,
this also ensures that direct-style \texttt{clone} will always be at least as powerful as \texttt{fork}.

We show more examples of direct-style process creation in section \ref{examples},
including use of process features which are difficult to use on Linux with \texttt{fork} or \texttt{posix\_spawn}.

Direct-style \texttt{clone} has acceptable performance cost,
and can outperform \texttt{fork} in some cases.
The performance overhead of creating processes with direct-style clone
is comparable to the performance cost of creating processes while inside a Linux namespace.
At \twosigma, we have deployed direct-style process creation in production
inside a library for distributed system deployment, Toplevel.
We evaluate in further detail in section \ref{evaluation}.
\section{Background on process-creation}\label{background}
As mentioned in the introduction,
on most operating systems today,
the available interfaces for process creation
are either "fork-style" or "spawn-style".
A few operating systems have a third style, which we refer to as "direct-style".

\subsection{Fork-style}
By far the most well-known fork-style interface is Unix's fork.\cite{forkhist}
When a parent process creates a child process by fork,
both processes execute concurrently,
and the child process typically inherits all but a few attributes from the parent process.\cite{tlpi}
The child process can call arbitrary system calls
to mutate the new process until it reaches the desired state,
typically ultimately calling \texttt{exec}.

Since the child process runs concurrently with the parent process,
communicating events in the child back to the parent requires some form of IPC,
complicating process creation.\cite{posix_spawn_error_pipe}
Even the basic interface of fork - one function call which returns twice - is unusual;
such functions are generally considered to be complex, advanced topics in other settings.
\cite{callcc_fork}\cite{continuations}\cite{bashcc}

A user of fork must think carefully about the characteristics of the process from which fork is being called.
Fork can have poor performance when called from processes with many memory mappings.\cite{forkroad}
This can be mitigated to some degree, but inflicts further complexity on the caller.\cite{vfork}\cite{vfork_dangerous}
Multi-threaded programs generally cannot use fork safely.\cite{forkroad}\cite{tlpi}
Some thread libraries provide partial mitigations for multi-threading issues,
but it's up to user code to make use of those mitigations.\cite{pthread_atfork}
\subsection{Spawn-style}
In a spawn-style interface,
details about the new process are provided up-front as arguments to some function
which creates the new process all at once.
Other details, such as security context, are inherited from the parent process.\cite{tlpi}\cite{create_process}
The most well known spawn-style interfaces are Unix's \texttt{posix\_spawn}
and Windows \texttt{CreateProcess};
the Unix shell also serves as a spawn-style process creation interface,
especially when used with Bernstein chaining.\cite{posix_spawn}\cite{create_process}\cite{chainloading}

Spawn-style APIs
can be called from any parent process without concerns about memory usage or multi-threading,
when used correctly.
\cite{forkroad}\cite{posix_spawn_error_pipe}

Unlike \texttt{fork}, a spawn-style interface provides an API for process creation
which is independent from existing system calls.\cite{spawn_h}\cite{subprocess_run}
Spawn-style interfaces therefore often lack support for some features,
and they also require learning a new interface distinct from the normal system call API,
both of which encourage systems programmers to avoid them.\cite{tlpi}\cite{posix_spawn_lwn}

Spawn-style interfaces, by their nature as a single call,
don't allow programs to branch on the results of individual modifications to the new process,
and have complex, or substantially less information, error reporting for such modifications.
\cite{posix_spawn}\cite{posix_spawn_error_pipe}
This further reduces the power of spawn-style interfaces.
\subsection{Direct-style}
A few academic operating systems, such as KeyKOS \cite{keykos}, seL4 \cite{sel4}
and some others \cite{exokernel} \cite{fuschia} \cite{singularity},
use another style of process creation.
We use the new term "direct-style" to refer to this,
since we know of no pre-existing concise and generic term for this style of process creation.

@@

However, we believe that a natural port of direct-style process creation to Linux
provides an interface that is just as simple to use as fork-style or spawn-style on Linux,
without their disadvantages.
Such an interface, like fork-style and spawn-style,
would implicitly copy some attributes to the new process,
rather than being capability-oriented and requiring all attributes to be specified explicitly.
This is the kind of direct-style interface that we contribute in this paper.
\section{Examples}\label{examples}
In this section,
we'll demonstrate direct-style process creation
by demonstrating several advanced uses of processes
using more sophisticated Linux features.
For concision, we assume in these examples that we are running with sufficient permissions in these examples,
but these examples can all be run without privileges with appropriate use of \verb|CLONE_NEWUSER|.

@@

\lstinputlisting[float=*ht,language=Python,label={fds},caption={Passing down FDs}]
{examples/fds/direct.py}
In Listing \ref{fds},
we first open a file with read-write permission in the local process.
Then we create a child process,
which inherits all file descriptors from its parent.
We indicate that we want to operate on the child's inherited copy of the file descriptor with \verb|child.inherit_fd|.
Then we disable \texttt{CLOEXEC} on the child's file descriptor,
which on Linux we can do by clearing the FD flags with \verb|fcntl(fd, F_SETFD, 0)|.
Finally, we execute some executable in the child process,
passing the file descriptor number as an argument;
since we disabled \texttt{CLOEXEC} on the file descriptor,
the executable will be able to use that file descriptor through that file descriptor number.
\subsection{Non-shared-memory concurrency}
Processes run concurrently,
which allows exploiting the parallelism of the underlying hardware.
Since processes don't share memory,
they can provide a less complex parallel programming environment
than shared-memory thread-based approaches.
